3D Magnetic modelling of ellipsoidal bodies

### 2017-02-08 manuscript/manuscript.tex
include the isotropic susceptibility case

Based on \citeauthor{maxwell1873}'s postulate, let us
assume that the body has a uniform magnetization given by
\begin{equation}
\mathbf{M} = \mathbf{K} \, \mathbf{H}^{\dagger} \: ,
\label{eq:M-KH}
\end{equation}
where $\mathbf{H}^{\dagger}$ is the resultant uniform magnetic field 
at any point within the body and $\mathbf{K}$ is a 
constant and symmetrical 2nd-order tensor representing the 
magnetic susceptibility of the body. 
In general, the susceptibility tensor $\mathbf{K}$ is represented,
in the main coordinate system, as follows:
\begin{equation}
\mathbf{K} = \mathbf{U}
\left[ \begin{array}{ccc}
k_{1} & 0 & 0 \\
0 & k_{2} & 0 \\
0 & 0 & k_{3} 
\end{array} \right] \mathbf{U}^{\top} \: ,
\label{eq:K}
\end{equation}
where $k_{1} > k_{2} > k_{3}$ are the 
\textit{principal susceptibilities} and $\mathbf{U}$ is
an orthogonal matrix whose columns $\mathbf{u}_{i}$,
$i = 1, 2, 3$, are unit vectors called 
\textit{principal directions} \citep{hrouda1982}.
For the particular case in which the principal directions 
coincide with the ellipsoid axes, the matrix $\mathbf{U}$ is
equal to the matrix $\mathbf{V}$ (Eqs. \ref{eq:V_triaxial_prolate}
and \ref{eq:V_oblate}), depending on the ellipsoid type.

@@

By considering $\tilde{\mathbf{r}}$ as a point within the
volume $\vartheta$ of the ellipsoid and using the Maxwell's postulate 
about the uniformity of the magnetic field $\mathbf{H}(\mathbf{r})$ 
inside ellipsoidal bodies, we can use Eq. \ref{eq:H-tilde}
for defining the resultant uniform magnetic field $\tilde{\mathbf{H}}^{\dagger}$
inside the ellipsoidal body as follows:
\begin{equation}
\tilde{\mathbf{H}}^{\dagger} = 
\left( \mathbf{I} - \tilde{\mathbf{N}}^{\dagger} \, \tilde{\mathbf{K}} \right)^{-1}
\tilde{\mathbf{H}}_{0} \: ,
\label{eq:Hi-tilde}
\end{equation}
where $\mathbf{I}$ is the identity matrix and 
$\tilde{\mathbf{N}}^{\dagger}$ represents the
transformed depolarization tensor evaluated within 
the volume $\vartheta$ of the ellipsoidal body. The elements
of $\tilde{\mathbf{N}}^{\dagger}$ will be presented in a 
latter section.

Let us pre-multiply the uniform internal field $\tilde{\mathbf{H}}^{\dagger}$
(Eq. \ref{eq:Hi-tilde}) by the transformed 
susceptibility tensor $\tilde{\mathbf{K}}$ 
(Eq. \ref{eq:H-tilde}) to obtain
\begin{equation}
\begin{split}
\tilde{\mathbf{M}} 
&= \tilde{\mathbf{K}} 
\left( \mathbf{I} - \tilde{\mathbf{N}}^{\dagger} \, \tilde{\mathbf{K}} \right)^{-1}
\tilde{\mathbf{H}}_{0} \\
&=  
\left( \mathbf{I} - \tilde{\mathbf{K}} \, \tilde{\mathbf{N}}^{\dagger} \right)^{-1}
\tilde{\mathbf{K}} \, \tilde{\mathbf{H}}_{0}
\end{split} \: ,
\label{eq:M-tilde}
\end{equation}
where $\tilde{\mathbf{M}}$
represents the transformed magnetization, as can be easily
verified by using Eqs. \ref{eq:M-KH} and \ref{eq:H-tilde}.
The matrix identity used for obtaining the second line of Eq. 
\ref{eq:M-tilde} is given by \citet[p. ~151]{searle1982}.

@@

FALAR QUE ESTA EQUACAO REPRESENTA O CASO HOMOGENEO ANISOTROPICO. 
FALAR QUE A MAIORIA DOS TRABALHOS DISCUTE O CASO HOMOGENEO ISOTROPICO
(CITAR TRABALHOS)
O CASO ANISOTROPICO FOI APRESENTADO PELA PRIMEIRA VEZ POR \citet{soliverez1981} and
posteriorly by \citet{emerson1985} and \citet{clark1986} (CHECAR).

FALAR DO CASO PARTICULAR EM QUE A SUSCETIBILIDADE EH MUITO FRACA E
A REMANENCIA EH NULA. NESTE CASO, A MAGNETIZACAO EH A FORMULA CLASSICA
DE MAGNETIZACAO INDUZIDA ENCONTRADA NOS LIVROS.

FALAR TAMBEM SOBRE O CASO ISOTROPICO E CITAR O TRABALHO DO GUO 2001.
USAR OS TERMOS INTRINSIC E EFFECTIVE MAGNETIZATIONS.

@@

The elements of the transformed depolarization tensor 
$\tilde{\mathbf{N}}^{\dagger}$ used to compute the 
uniform magnetic field $\tilde{\mathbf{H}}^{\dagger}$
(Eq. \ref{eq:Hi-tilde}) and the magnetization $\tilde{\mathbf{M}}$ 
(Eqs. \ref{eq:M-tilde} and \ref{eq:M-tilde-remanence})
are calculated according to Eq. \ref{eq:nij-tilde}, with 
$\tilde{f}(\tilde{\mathbf{r}})$ given by 
$\tilde{f}^{\dagger}(\tilde{\mathbf{r}})$ (Eq. \ref{eq:fi-tilde}).
As we have already pointed out, the  
$\tilde{f}^{\dagger}(\tilde{\mathbf{r}})$ (Eq. \ref{eq:fi-tilde}) is a
quadratic function of the spatial coordinates $\tilde{x}$,
$\tilde{y}$ and $\tilde{z}$. Consequently, the 
elements $\tilde{n}^{\dagger}_{ij}$, $i = 1, 2, 3$, 
$j = 1, 2, 3$, of $\tilde{\mathbf{N}}^{\dagger}$ 
do not depend on the transformed position
vector $\tilde{\mathbf{r}}$ (equation \ref{eq:coord_transformation}). 
Besides, the off-diagonal elements are zero and
the diagonal elements are given by \citep{stoner1945}:
\begin{equation}
\tilde{n}^{\dagger}_{ii} = \frac{abc}{2}
\int_{0}^{\infty} \frac{1}{\left( e_{i}^{2} 
+ u \right) R(u)} \, du \: , \quad i = 1, 2, 3 \: ,
\label{eq:n-tilde-dagger-ii}
\end{equation}
where $R(u)$ is defined by Eq. \ref{eq:R} and
$e_{1} = a$, $e_{2} = b$, and $e_{3} = c$. These elements 
are commonly known as \textit{demagnetizing factors}
and are defined according to the ellipsoid type.
Notice that, according to Eqs. \ref{eq:H-tilde} 
and \ref{eq:N-tilde-VT-N-V},
\begin{equation}
\mathbf{N}(\mathbf{r}) = 
\mathbf{V} \, \tilde{\mathbf{N}}^{\dagger} \, 
\mathbf{V}^{\top} \: ,
\label{eq:N-V-N-dagger-VT}
\end{equation}
where $\tilde{\mathbf{N}}^{\dagger}$ (Eqs. \ref{eq:Hi-tilde}, 
\ref{eq:M-tilde} and \ref{eq:M-tilde-remanence}) is a diagonal 
matrix and $\mathbf{V}$ (Eqs. \ref{eq:V_triaxial_prolate} and
\ref{eq:V_oblate}) is an orthogonal matrix.
This equation shows that, for the particular case 
in which $\mathbf{r}$ and consequently $\tilde{\mathbf{r}}$
represent a point inside the
volume $\vartheta$ of the ellipsoid, the elements 
$\tilde{n}^{\dagger}_{ii}$ (Eq. \ref{eq:n-tilde-dagger-ii})
of $\tilde{\mathbf{N}}^{\dagger}$ represent the eigenvalues while the 
columns of $\mathbf{V}$ represent the eigenvectors of the 
original depolarization tensor $\mathbf{N}(\mathbf{r})$.

@@

The elements $\tilde{n}^{\ddagger}_{ij}(\tilde{\mathbf{r}})$,
$i = 1, 2, 3$, $j = 1, 2, 3$, of the transformed depolarization tensor 
$\tilde{\mathbf{N}}^{\ddagger}(\tilde{\mathbf{r}})$ used to compute
the magnetic field $\Delta \tilde{\mathbf{H}}(\tilde{\mathbf{r}})$ 
(Eq. \ref{eq:delta-H-tilde}) and
the total-field anomaly $\Delta \tilde{T}(\tilde{\mathbf{r}})$
(Eqs. \ref{eq:delta-T-tilde} and \ref{eq:delta-T-tilde-approx}) 
are calculated according to Eq. \ref{eq:nij-tilde}, with 
$\tilde{f}(\tilde{\mathbf{r}})$ given by $\tilde{f}^{\ddagger}(\tilde{\mathbf{r}})$
(Eq. \ref{eq:fe-tilde}).
By following \citet{clark1986}, the diagonal elements $\tilde{n}^{\ddagger}_{ii}$
and the off-diagonal elements $\tilde{n}^{\ddagger}_{ij}$, $i = 1, 2, 3$,
$j = 1, 2, 3$, are given by
\begin{equation}
\tilde{n}^{\ddagger}_{ii}(\tilde{\mathbf{r}}) =
-\frac{abc}{2}
\left( \frac{\partial \lambda}{\partial \tilde{r}_{i}} \, h_{i} \, \tilde{r}_{i}
+ g_{i} \right)
\label{eq:n-tilde-dagger-ii}
\end{equation}
and
\begin{equation}
\tilde{n}^{\ddagger}_{ij}(\tilde{\mathbf{r}}) =
-\frac{abc}{2} \left(
\frac{\partial \lambda}{\partial \tilde{r}_{i}} \, h_{j} \, \tilde{r}_{j} 
\right) \: ,
\label{eq:n-tilde-dagger-ij}
\end{equation}
where
\begin{equation}
h_{i} = -\frac{1}{\left( e_{i}^{2} + \lambda \right) R(\lambda)} \: ,
\label{eq:hi}
\end{equation}
\begin{equation}
g_{i} = \int_{\lambda}^{\infty} \frac{1}{\left( e_{i}^{2} + u \right) R(u)} du \: ,
\label{eq:gi}
\end{equation}
$e_{1} = a$, $e_{2} = b$, $e_{3} = c$, and 
$\frac{\partial \lambda}{\partial \tilde{r}_{i}}$
is defined in Appendix B (Eq. \ref{eq:dlambda}).
The functions $g_{i}$ (Eq. \ref{eq:gi}) are defined according
to the ellipsoid type.

@@

@@

\conclusions  %% \conclusions[modified heading if necessary]
TEXT

>>>>

Based on \citeauthor{maxwell1873}'s postulate, let us
assume that the body has a uniform magnetization given by
\begin{equation}
\mathbf{M} = \mathbf{K} \, \mathbf{H}^{\dagger} \: ,
\label{eq:M-KH}
\end{equation}
where $\mathbf{H}^{\dagger}$ is the resultant uniform magnetic field 
at any point within the body and $\mathbf{K}$ is a 
constant and symmetrical 2nd-order tensor representing the 
magnetic susceptibility of the body. 
In general, the susceptibility tensor $\mathbf{K}$ is represented,
in the main coordinate system, as follows:
\begin{equation}
\mathbf{K} = \mathbf{U}
\left[ \begin{array}{ccc}
k_{1} & 0 & 0 \\
0 & k_{2} & 0 \\
0 & 0 & k_{3} 
\end{array} \right] \mathbf{U}^{\top} \: ,
\label{eq:K}
\end{equation}
where $k_{1} > k_{2} > k_{3}$ are the 
\textit{principal susceptibilities} and $\mathbf{U}$ is
an orthogonal matrix whose columns $\mathbf{u}_{i}$,
$i = 1, 2, 3$, are unit vectors called 
\textit{principal directions} \citep{hrouda1982}.
For the particular case in which the principal directions 
coincide with the ellipsoid axes, the matrix $\mathbf{U}$ is
equal to the matrix $\mathbf{V}$ (Eqs. \ref{eq:V_triaxial_prolate}
and \ref{eq:V_oblate}), depending on the ellipsoid type.
Another important particular case is that in which the
susceptibility is isotropic and, consequently, the principal
susceptibilities $k_{1}$, $k_{2}$, and $k_{3}$ (Eq. \ref{eq:K})
are equal to a constant $\chi$. In this case, the susceptibility
tensor $\mathbf{K}$ (Eq. \ref{eq:K}) assumes the particular form
\begin{equation}
\mathbf{K} = \chi \, \mathbf{I} \: ,
\label{eq:K-isotropic}
\end{equation}
where $\mathbf{I}$ represents the identity matrix.

@@

By considering $\tilde{\mathbf{r}}$ as a point within the
volume $\vartheta$ of the ellipsoid and using the Maxwell's postulate 
about the uniformity of the magnetic field $\mathbf{H}(\mathbf{r})$ 
inside ellipsoidal bodies, we can use Eq. \ref{eq:H-tilde}
for defining the resultant uniform magnetic field $\tilde{\mathbf{H}}^{\dagger}$
inside the ellipsoidal body as follows:
\begin{equation}
\tilde{\mathbf{H}}^{\dagger} = 
\left( \mathbf{I} - \tilde{\mathbf{N}}^{\dagger} \, \tilde{\mathbf{K}} \right)^{-1}
\tilde{\mathbf{H}}_{0} \: ,
\label{eq:Hi-tilde}
\end{equation}
where $\mathbf{I}$ is the identity matrix and 
$\tilde{\mathbf{N}}^{\dagger}$ is a diagonal matrix
representing the transformed depolarization tensor 
evaluated within the volume $\vartheta$ of the ellipsoidal body. 
The elements of $\tilde{\mathbf{N}}^{\dagger}$ will be 
presented in a latter section.

Let us pre-multiply the uniform internal field $\tilde{\mathbf{H}}^{\dagger}$
(Eq. \ref{eq:Hi-tilde}) by the transformed 
susceptibility tensor $\tilde{\mathbf{K}}$ 
(Eq. \ref{eq:H-tilde}) to obtain
\begin{equation}
\begin{split}
\tilde{\mathbf{M}} 
&= \tilde{\mathbf{K}} 
\left( \mathbf{I} - \tilde{\mathbf{N}}^{\dagger} \, \tilde{\mathbf{K}} \right)^{-1}
\tilde{\mathbf{H}}_{0} \\
&=  
\left( \mathbf{I} - \tilde{\mathbf{K}} \, \tilde{\mathbf{N}}^{\dagger} \right)^{-1}
\tilde{\mathbf{K}} \, \tilde{\mathbf{H}}_{0}
\end{split} \: ,
\label{eq:M-tilde}
\end{equation}
where $\tilde{\mathbf{M}}$
represents the transformed magnetization, as can be easily
verified by using Eqs. \ref{eq:M-KH} and \ref{eq:H-tilde}.
The matrix identity used for obtaining the second line of Eq. 
\ref{eq:M-tilde} is given by \citet[p. ~151]{searle1982}.
For the particular case in which the susceptibility is isotropic
(Eq. \ref{eq:K-isotropic}) and $\chi << 1$, the magnetization
$\tilde{\mathbf{M}}$ (Eq. \ref{eq:M-tilde}) can be approximated
as follows:
\begin{equation}
\tilde{\mathbf{M}} \approx \chi \, \tilde{\mathbf{H}}_{0} \: ,
\label{eq:M-tilde-chi-H0}
\end{equation}
which is the classical equation describing the induced
magnetization in applied geophysics \citep[p. 87]{blakely1996}.
Notice that, in this particular case, the magnetization is
parallel to the inducing field $\tilde{\mathbf{H}}_{0}$.

FALAR DO CASO EM QUE A INDUCAO EH PARALELA A UM EIXO DO ELIPSOIDE

@@

@@

The elements of the transformed depolarization tensor 
$\tilde{\mathbf{N}}^{\dagger}$ used to compute the 
uniform magnetic field $\tilde{\mathbf{H}}^{\dagger}$
(Eq. \ref{eq:Hi-tilde}) and the magnetization $\tilde{\mathbf{M}}$ 
(Eqs. \ref{eq:M-tilde} and \ref{eq:M-tilde-remanence})
are calculated according to Eq. \ref{eq:nij-tilde}, with 
$\tilde{f}(\tilde{\mathbf{r}})$ given by 
$\tilde{f}^{\dagger}(\tilde{\mathbf{r}})$ (Eq. \ref{eq:fi-tilde}).
As we have already pointed out, the  
$\tilde{f}^{\dagger}(\tilde{\mathbf{r}})$ (Eq. \ref{eq:fi-tilde}) is a
quadratic function of the spatial coordinates $\tilde{x}$,
$\tilde{y}$ and $\tilde{z}$. Consequently, the 
elements $\tilde{n}^{\dagger}_{ij}$, $i = 1, 2, 3$, 
$j = 1, 2, 3$, of $\tilde{\mathbf{N}}^{\dagger}$ 
do not depend on the transformed position
vector $\tilde{\mathbf{r}}$ (equation \ref{eq:coord_transformation}). 
Besides, the off-diagonal elements are zero and
the diagonal elements are given by \citep{stoner1945}:
\begin{equation}
\tilde{n}^{\dagger}_{ii} = \frac{abc}{2}
\int_{0}^{\infty} \frac{1}{\left( e_{i}^{2} 
+ u \right) R(u)} \, du \: , \quad i = 1, 2, 3 \: ,
\label{eq:n-tilde-dagger-ii}
\end{equation}
where $R(u)$ is defined by Eq. \ref{eq:R} and
$e_{1} = a$, $e_{2} = b$, and $e_{3} = c$. These elements 
are commonly known as \textit{demagnetizing factors}
and are defined according to the ellipsoid type.
Here, we calculate the demagnetizing factors in the SI
system. Consequently, they satisfies the condition
$\tilde{n}^{\dagger}_{11} + \tilde{n}^{\dagger}_{22} + 
\tilde{n}^{\dagger}_{33} = 1$, independently of the
ellipsoid type.
Notice that, according to Eqs. \ref{eq:H-tilde} 
and \ref{eq:N-tilde-VT-N-V},
\begin{equation}
\mathbf{N}(\mathbf{r}) = 
\mathbf{V} \, \tilde{\mathbf{N}}^{\dagger} \, 
\mathbf{V}^{\top} \: ,
\label{eq:N-V-N-dagger-VT}
\end{equation}
where $\tilde{\mathbf{N}}^{\dagger}$ (Eqs. \ref{eq:Hi-tilde}, 
\ref{eq:M-tilde} and \ref{eq:M-tilde-remanence}) is a diagonal 
matrix and $\mathbf{V}$ (Eqs. \ref{eq:V_triaxial_prolate} and
\ref{eq:V_oblate}) is an orthogonal matrix.
This equation shows that, for the particular case 
in which $\mathbf{r}$ and consequently $\tilde{\mathbf{r}}$
represent a point inside the
volume $\vartheta$ of the ellipsoid, the elements 
$\tilde{n}^{\dagger}_{ii}$ (Eq. \ref{eq:n-tilde-dagger-ii})
of $\tilde{\mathbf{N}}^{\dagger}$ represent the eigenvalues while the 
columns of $\mathbf{V}$ represent the eigenvectors of the 
original depolarization tensor $\mathbf{N}(\mathbf{r})$.

@@

The elements $\tilde{n}^{\ddagger}_{ij}(\tilde{\mathbf{r}})$,
$i = 1, 2, 3$, $j = 1, 2, 3$, of the transformed depolarization tensor 
$\tilde{\mathbf{N}}^{\ddagger}(\tilde{\mathbf{r}})$ used to compute
the magnetic field $\Delta \tilde{\mathbf{H}}(\tilde{\mathbf{r}})$ 
(Eq. \ref{eq:delta-H-tilde}) and
the total-field anomaly $\Delta \tilde{T}(\tilde{\mathbf{r}})$
(Eqs. \ref{eq:delta-T-tilde} and \ref{eq:delta-T-tilde-approx}) 
are calculated according to Eq. \ref{eq:nij-tilde}, with 
$\tilde{f}(\tilde{\mathbf{r}})$ given by $\tilde{f}^{\ddagger}(\tilde{\mathbf{r}})$
(Eq. \ref{eq:fe-tilde}).
By following \citet{clark1986}, the diagonal elements $\tilde{n}^{\ddagger}_{ii}$
and the off-diagonal elements $\tilde{n}^{\ddagger}_{ij}$, $i = 1, 2, 3$,
$j = 1, 2, 3$, are given by
\begin{equation}
\tilde{n}^{\ddagger}_{ii}(\tilde{\mathbf{r}}) =
-\frac{abc}{2}
\left( \frac{\partial \lambda}{\partial \tilde{r}_{i}} \, h_{i} \, \tilde{r}_{i}
+ g_{i} \right)
\label{eq:n-tilde-dagger-ii}
\end{equation}
and
\begin{equation}
\tilde{n}^{\ddagger}_{ij}(\tilde{\mathbf{r}}) =
-\frac{abc}{2} \left(
\frac{\partial \lambda}{\partial \tilde{r}_{i}} \, h_{j} \, \tilde{r}_{j} 
\right) \: ,
\label{eq:n-tilde-dagger-ij}
\end{equation}
where
\begin{equation}
h_{i} = -\frac{1}{\left( e_{i}^{2} + \lambda \right) R(\lambda)} \: ,
\label{eq:hi}
\end{equation}
\begin{equation}
g_{i} = \int_{\lambda}^{\infty} \frac{1}{\left( e_{i}^{2} + u \right) R(u)} du \: ,
\label{eq:gi}
\end{equation}
$e_{1} = a$, $e_{2} = b$, $e_{3} = c$, and 
$\frac{\partial \lambda}{\partial \tilde{r}_{i}}$
is defined in Appendix B (Eq. \ref{eq:dlambda}).
The functions $g_{i}$ (Eq. \ref{eq:gi}) are defined according
to the ellipsoid type.
Notice that the elements $\tilde{n}^{\ddagger}_{ij}$
(Eqs. \ref{eq:n-tilde-dagger-ii} and \ref{eq:n-tilde-dagger-ij}) 
are proportional to the second derivatives of the function
$\tilde{f}^{\ddagger}(\tilde{\mathbf{r}})$ (Eq. \ref{eq:fe-tilde}),
which is harmonic. Consequently, the diagonal elements defined
in Eq. \ref{eq:n-tilde-dagger-ii} satisfies the condition
$\tilde{n}^{\ddagger}_{11}(\tilde{\mathbf{r}}) +
\tilde{n}^{\ddagger}_{22}(\tilde{\mathbf{r}}) +
\tilde{n}^{\ddagger}_{33}(\tilde{\mathbf{r}}) = 0$ for any
point $\tilde{\mathbf{r}}$ outside the ellipsoid, independently
of the ellipsoid type.

@@

\section{Computational implementation and testing}

The code is implemented in Python language,
by using use the NumPy and SciPy libraries \citep{vanderwalt2011},
as part of the open-source source library
Fatiando a Terra \citep{uieda2013}.
Because of that, it is very modular and has a
test suite formed by a considerable number of 
assertions, unit tests, doc tests and integration tests.
For readers interested in automated code testing,
we recommend taking a look at the 
freely reusable lessons of the Software Carpentry 
(\url{https://software-carpentry.org/}) and at the
description of the doctest module that is available
in the documentation for Python 2.X 
(\url{https://docs.python.org/2/library/doctest.html}).

@@

\section{Conclusions}  %% \conclusions[modified heading if necessary]
TEXT
